\section{Related Work}
\label{sec:related}

CRAF-X sits at the intersection of four research threads: (i) deep
camera--LiDAR fusion for 3D object detection, (ii) adversarial attacks on
autonomous-vehicle (AV) sensors, (iii) robust and certified defenses for
single- and multi-modal perception, and (iv) explainable AI (XAI) for
safety-critical driving systems. We review each in turn, positioning
CRAF-X as the first framework that unifies cross-modal consistency
probing, spatially dynamic gated fusion, and auditable Trust Maps in a
single detector.

\subsection{Deep camera--LiDAR fusion for 3D object detection}

Camera--LiDAR fusion has matured from early proposal-level schemes to
transformer-based dense fusion. Early approaches such as
AVOD~\cite{KuAVOD2018} and Frustum PointNets~\cite{QiFrustum2018}
combined 2D region proposals with LiDAR geometry but are brittle to
failures in the image branch that propagate downstream. Sequential
point-level fusion was popularised by PointPainting~\cite{VoraPointPainting2020},
which decorates raw LiDAR points with per-pixel segmentation logits, and
later refined by EPNet~\cite{HuangEPNet2020} with per-point semantic
enhancement and a consistency loss. Multimodal Virtual
Points~\cite{YinMVP2021} go further, using 2D instance masks to
densify sparse LiDAR; yet all three pipelines inherit the camera
branch's adversarial surface.
Feature-level BEV fusion is now the dominant paradigm:
BEVFusion~\cite{LiuBEVFusion2023} projects both modalities into a shared
bird's-eye-view tensor, while the concurrent
BEVFusion~\cite{LiangBEVFusion2022} deliberately decouples the streams
to ensure graceful degradation when LiDAR is absent.
TransFusion~\cite{BaiTransFusion2022},
DeepFusion~\cite{LiDeepFusion2022},
AutoAlignV2~\cite{ChenAutoAlignV22022}, and the Cross Modal
Transformer~\cite{YanCMT2023} instead adopt attention: object queries
softly attend to image tokens, improving tolerance to calibration noise
and missing modalities. Feng~et~al.'s survey~\cite{FengSurvey2021}
frames these methods along a \emph{what/when/how-to-fuse} taxonomy.
CRAF-X builds on BEV-attention fusion but departs from prior work in
treating every cross-modal association not as a soft prior but as a
hypothesis to be \emph{probed} against an adversarially robust
consistency criterion before contributing to the output.

Sgaravatti~et~al.~\cite{SgaravattiLCF3D2026} very recently proposed LCF3D,
a late-cascade fusion framework that explicitly crossvalidates LiDAR
3D proposals against RGB 2D detections to suppress false positives and
recover false negatives, achieving strong domain-generalisation on KITTI
and nuScenes. Their cross-modal validation logic directly motivates the
coordinate-level consistency probe in CRAF-X, extending it to an
adversarially hardened setting.

\subsection{Adversarial attacks on AV perception}

AV perception has been attacked at every level of the stack. On the
camera branch, Eykholt~et~al.'s RP2 stop-sign
attack~\cite{EykholtRP22018} and Thys~et~al.'s person-hiding
patch~\cite{ThysPatch2019} established that small physical perturbations
can reliably fool image detectors in the wild. LiDAR-side attacks have
progressed from spoofed laser injections against Apollo perception
(Cao~et~al.~\cite{CaoAdvLiDAR2019}) to generalised black-box sensor
attacks~\cite{SunUSENIX2020}, physically realisable universal adversarial
meshes that hide vehicles~\cite{TuPhysical2020}, and most recently
PLA-LiDAR~\cite{JinPLALiDAR2023}, which injects thousands of
controlled points with a custom laser transceiver. These single-modality
attacks motivate the hope that fusion is inherently robust.
That hope has been dispelled by three pivotal works. MSF-ADV~\cite{CaoMSFADV2021}
jointly optimises a physical object to fool both image and LiDAR
branches of production fusion stacks with over 90\% attack success. The
frustum attack of Hallyburton~et~al.~\cite{HallyburtonFrustum2022}
designs spoofed LiDAR points that \emph{preserve} cross-modal
consistency, evading na\"{i}ve geometric sanity checks. Most recently,
Cheng~et~al.'s ``Fusion is Not Enough''~\cite{ChengFusion2024} showed
that camera-only adversarial patches are sufficient to collapse six
state-of-the-art fusion detectors, because static fusion weights allow
the weakest modality to dominate. Together these results expose two
gaps that CRAF-X addresses directly: consistency checks must look beyond
coincident projection, and fusion weights must be input-adaptive and
\emph{spatially local} rather than global.

Sahu~et~al.~\cite{SahuAD22026} conducted a 2026 closed-loop evaluation
of end-to-end AV agents under three attack vectors---acoustic-wave blur,
electromagnetic image distortion, and ghost-object perturbations---in the
CARLA simulator, observing driving-score collapses of up to 99\% and
proposing a lightweight attention-based detector (AD$^2$) for run-time
threat identification. Their findings reinforce the urgency of in-model
detection mechanisms such as CRAF-X's consistency probe.

\subsection{Robust and certified defenses for multi-modal perception}

Defensive research has largely treated each modality in isolation.
On the image side, Madry~et~al.'s PGD adversarial
training~\cite{MadryPGD2018} remains the canonical empirical defense,
while randomised smoothing~\cite{CohenSmoothing2019} and
PixelDP~\cite{LecuyerPixelDP2019} provide certified $\ell_2$ and
differential-privacy-based guarantees at ImageNet scale. For point
clouds, PointGuard~\cite{LiuPointGuard2021} offers the first provable
robustness against point addition, deletion, and modification via
random sub-sampling and majority vote, and DUP-Net~\cite{ZhouDUPNet2019}
exemplifies the purification-then-classify paradigm using statistical
outlier removal and a learned upsampler.
Multi-modal defenses are comparatively scarce. The closest precedent is
Tu~et~al.~\cite{TuMultiSensor2021}, who demonstrate physically realisable
universal adversaries against fusion detectors and show that
adversarial training with feature denoising improves robustness, while
explicitly noting that the image branch remains the weaker link.
AdvSim~\cite{WangAdvSim2021} complements per-frame defenses with
scenario-level stress testing by resynthesising LiDAR under adversarial
actor trajectories. None of these works, however, perform
\emph{cross-modal} consistency probing or allow fine-grained
quarantining of corrupted regions. CRAF-X closes this gap by combining
coordinate-level agreement tests with PGD-style adversarial training of
the fusion head, and empirically benchmarks against both
Tu~et~al.~\cite{TuMultiSensor2021} and the attacks of
\cite{CaoMSFADV2021,HallyburtonFrustum2022,ChengFusion2024}.

\subsection{Dynamic and uncertainty-aware fusion}

The idea of letting the network decide \emph{how much} to trust each
modality has a long history. Arevalo~et~al.'s Gated Multimodal
Unit~\cite{ArevaloGMU2017} introduced per-sample sigmoid gates, while
the Sparsely-Gated Mixture-of-Experts
layer~\cite{ShazeerMoE2017} generalised this to routing across
sub-networks. In robotics, Mees~et~al.~\cite{MeesIROS2016} first showed
that adaptive weights across RGB, depth, and motion experts improve
object detection under changing lighting and motion blur; in AV
perception, Kim~et~al.'s Gated Information Fusion
network~\cite{KimGIF2018} and Valada~et~al.'s
SSMA~\cite{ValadaSSMA2020} introduced spatially local gating across
sensor streams, explicitly designed to degrade gracefully when one
modality fails. Bijelic~et~al.~\cite{BijelicFog2020} extended this with
entropy-steered adaptive fusion for unseen adverse weather, while
HydraFusion~\cite{MalawadeHydra2022} selects among early, late, and
intermediate fusion branches based on driving context.
Closely related is uncertainty-aware fusion: Kendall, Gal, and
Cipolla~\cite{KendallMTL2018} weight multi-task losses by homoscedastic
uncertainty, and CLOCs~\cite{PangCLOCs2020} performs learned
object-candidate fusion before NMS.

Recent work has tightened the coupling between dynamic gating and
explicit, queryable reliability signals. Sadeghian~et~al.'s
ReliFusion~\cite{SadeghianReliFusion2025} employs cross-modality
contrastive learning to assign per-modality confidence scores in the
BEV space, feeding a Confidence-Weighted Mutual Cross-Attention
mechanism that gracefully re-weights LiDAR and camera contributions
under limited-FOV LiDAR or sensor-malfunction scenarios on nuScenes.
Chen~et~al.'s HyperDUM~\cite{ChenHyperDUM2025}, presented at CVPR 2025,
takes a complementary approach: hyperdimensional projection is used to
compute channel- and patch-wise epistemic uncertainty with a single
forward pass, producing spatial uncertainty maps that reweight modality
features before fusion across camera, LiDAR, and event-camera streams
under fog, rain, and sun corruptions on the DeLiVER and aiMotive benchmarks,
with substantially fewer parameters than Monte-Carlo baselines.

CRAF-X differs from this line in two key respects. First, our gating
criterion is driven by a \emph{probed cross-modal disagreement signal},
not by static input statistics or entropy; this allows it to detect
targeted adversaries that pass conventional confidence filters.
Second, our gating is applied at BEV-pixel granularity so that a
localised adversarial patch triggers a localised quarantine rather than
demoting the entire modality. Crucially, unlike ReliFusion and HyperDUM,
our gating signal is simultaneously a \emph{spatially-grounded
explainability output}---the Modal Attribution Map---rather than an
internal weighting coefficient that remains opaque to downstream auditors.

\subsection{Explainable AI for Safety Auditing: BEV Modal Attribution and Trust Maps}

Trust Maps sit within a broader literature on post-hoc explanation.
Gradient-based saliency methods---Grad-CAM~\cite{SelvarajuGradCAM2017},
SHAP~\cite{LundbergSHAP2017}, LIME~\cite{RibeiroLIME2016}, and
Integrated Gradients~\cite{SundararajanIG2017}---are the standard tools
for attributing a 2D prediction back to input pixels, and have
recognised 3D analogues: Zheng~et~al.'s PointCloud
saliency maps~\cite{ZhengPCSaliency2019} extend the idea to point-wise
attribution.
In the AV domain, Kim and Canny~\cite{KimCausal2017} pioneered causal
visual attention for end-to-end driving, and Zablocki~et~al.'s
survey~\cite{ZablockiSurvey2022} and Omeiza~et~al.'s
survey~\cite{OmeizaSurvey2022} catalogue the subsequent explosion of
XAI methods for vision-based AV and their regulatory relevance. A
prerequisite for trustworthy explanations is well-calibrated
confidence, studied comprehensively by Guo~et~al.~\cite{GuoCalibration2017}.

Two 2025 surveys sharpen the framing for multi-modal XAI.
Yeong~et~al.~\cite{YeongXAIAV2025} systematically review the intersection
of multi-sensor fusion (camera, LiDAR, radar, GPS) with post-hoc and
intrinsic XAI, identifying the absence of \emph{real-time, stakeholder-aware}
explanations as a primary barrier to AV deployment and certification.
A concurrent Sensors review~\cite{SensorsMSFReview2025} surveying BEV-centric
and token-level cross-modal fusion explicitly flags ``limited interpretability''
as one of three primary deployment blockers alongside spatio-temporal
misalignment and domain shift, and calls for Shapley-style sensor-wise
contribution maps and attention-rollout methods as emerging solutions.

Despite this activity, explanations for \emph{multi-modal} AV perception
remain largely unstudied: prior work rarely attributes a 3D detection back
to the responsible modality, let alone at spatial granularity.
Park~et~al.'s Layer-wise Modality Decomposition
(LMD)~\cite{ParkLMD2025}, presented at NeurIPS 2025, addresses this
gap directly and represents the closest published precedent to CRAF-X's
Trust Maps. LMD propagates modality-specific contributions separately
through every non-linear layer of a pretrained fusion network---activations,
LayerNorm, attention---using a first-order Taylor / layer-wise relevance
propagation linearisation with a ``ratio rule'' that preserves model
behaviour. The result is a per-layer, \emph{post-hoc} attribution map for
each input modality, validated on BEVFusion and RCBEVDet across camera,
LiDAR, and radar streams on nuScenes. While LMD is the first method to
attribute 3D fusion predictions to individual modalities, it remains a
\emph{post-hoc} tool requiring layer-by-layer linearisation of a frozen
model. CRAF-X, by contrast, produces Trust Maps as a \emph{native
architectural output} co-trained with the detection objective, yielding
attributions that are causally aligned with the model's live gating
decisions rather than being reconstructed from gradients after the fact.

Zeng~et~al.'s RESAR-BEV~\cite{ZengRESARBEV2025} demonstrates that
interpretability can alternatively be designed \emph{intrinsically} into
a fusion pipeline. Their progressive residual autoregressive architecture
for camera--radar BEV segmentation decomposes predictions into
coarse-to-fine stages---early stages establish road topology, later stages
refine lane boundaries and small objects---providing stage-level error
localisation on nuScenes at competitive accuracy and real-time speed.
RESAR-BEV confirms that architectural interpretability and fusion
performance are not at odds; CRAF-X extends this philosophy into the
adversarially hardened camera--LiDAR 3D detection setting, where the
consistency probe replaces the stage-decomposition mechanism.

The XAI-for-AV literature has also explored explanation modalities
beyond spatial maps. Zarghani~et~al.~\cite{ZarghaniExplainableAD2025}
propose a multimodal framework that fuses video (VideoMAE), sensor data,
and natural-language text (BERT) to jointly predict driving actions and
generate textual rationales, achieving a BLEU-4 score of 0.75 on BDD-X
while fostering regulatory compliance through human-readable explanations.
While orthogonal to CRAF-X's spatial attribution focus, this line of work
illustrates the breadth of the emerging ``explainability-by-design''
programme for AV perception.

CRAF-X's Trust Maps fill the gap identified across all of these works
by exposing, for every BEV cell, which modality the fusion gate trusted
and with what calibrated confidence---and doing so \emph{natively}, as a
byproduct of the gating process rather than a separate post-hoc step.
In combination with the consistency probe, this provides safety auditors
with a falsifiable record of why each object was accepted or rejected---an
artefact that, we argue, is as important as the detection itself for the
certification of adversarially robust AV perception.

Almadhor~et~al.~\cite{AlmadhorMSF2026} integrated LiDAR, radar, and
camera fusion with a Transformer backbone and an explicit XAI module on
nuScenes, demonstrating that embedding explainability into the perception
pipeline---rather than appending it post hoc---yields more faithful
spatial attributions; this motivates CRAF-X's design of Trust Maps as a
first-class architectural output rather than a visualisation afterthought.
Similarly, Wei~et~al.~\cite{WeiSafeDrive2026} proposed SafeDrive-Fusion,
an illumination-adaptive dynamic attention framework that achieves 87.5\%
accuracy under extreme low-light by learning spatially varying modality
weights---an orthogonal robustness axis to adversarial corruption that
future extensions of CRAF-X may incorporate.